---
title: "Working with aggregate functions"
author: 
  - name: Giulia Pais
    affiliation: | 
     San Raffaele Telethon Institute for Gene Therapy - SR-Tiget, 
     Via Olgettina 60, 20132 Milano - Italia
    email: giuliapais1@gmail.com, calabria.andrea@hsr.it
output: 
  BiocStyle::html_document:
    self_contained: yes
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
date: "`r doc_date()`"
package: "`r pkg_ver('ISAnalytics')`"
vignette: >
  %\VignetteIndexEntry{Working with aggregate functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    crop = NULL
    ## Related to
    ## https://stat.ethz.ch/pipermail/bioc-devel/2020-April/016656.html
)
```


```{r vignetteSetup, echo=FALSE, message=FALSE, warning = FALSE}
## Track time spent on making the vignette
startTime <- Sys.time()

## Bib setup
library("knitcitations")

## Load knitcitations with a clean bibliography
cleanbib()
cite_options(hyperlink = "to.doc", citation_format = "text", style = "html")

## Write bibliography information
bib <- c(
    R = citation(),
    BiocStyle = citation("BiocStyle")[1],
    knitcitations = citation("knitcitations")[1],
    knitr = citation("knitr")[1],
    rmarkdown = citation("rmarkdown")[1],
    sessioninfo = citation("sessioninfo")[1],
    testthat = citation("testthat")[1],
    ISAnalytics = citation("ISAnalytics")[1]
)

write.bibtex(bib, file = "aggregate_function_usage.bib")
```

# Introduction

In this vignette we're going to explain in detail how to use functions of the 
aggregate family, namely:

1. `aggregate_metadata`
2. `aggregate_values_by_key`

## How to install ISAnalytics

To install the package run the following code:

```{r installBioc, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
      install.packages("BiocManager")
  }
BiocManager::install("ISAnalytics")
```

To install from GitHub:

```{r installGitHub, eval=FALSE}
if (!require(devtools)) {
    install.packages("devtools")
}
devtools::install_github("calabrialab/ISAnalytics", build_vignettes = TRUE)
```
```{r}
library(ISAnalytics)
```

## Setting options

`ISAnalytics` has a verbose option that allows some functions to print 
additional information to the console while they're executing. 
To disable this feature do:

```{r OptVerbose, eval=FALSE}
# DISABLE
options("ISAnalytics.verbose" = FALSE)

# ENABLE
options("ISAnalytics.verbose" = TRUE)

```

Some functions also produce report in a user-friendly HTML format, 
to set this feature:

```{r OptWidg, eval=FALSE}
# DISABLE HTML REPORTS
options("ISAnalytics.widgets" = FALSE)

# ENABLE HTML REPORTS
options("ISAnalytics.widgets" = TRUE)
```

# Aggregating metadata

We refer to information contained in the association file as "metadata": 
sometimes it's useful to obtain collective information based on a certain 
group of variables we're interested in. The function `aggregate_metadata` 
does just that, according to the grouping variables, meaning the names of 
the columns in the association file to perform a `group_by` operation with, 
creates a summary which includes:

* `FusionPrimerPCRDate` - The minimum date in the group for this variable
* `LinearPCRDate` - The minimum date in the group for this variable
* `VCN` - The mean of "VCN" column for each group
* `Avg_DNAngUsed` - The mean of "DNAngUsed "column for each group
* `Kapa` - The mean of "Kapa" column for each group
* `DNAngUsed` - The sum of "DNAngUsed" column for each group
* `ulForPool` - The sum of "ulForPool" column for each group
* `AggregateMeta` - A string obtained by concatenation of all of the grouping 
variables separated by "_"

## Typical workflow

Import the association file via `import_assocition_file`. If you need more 
information on import function please view the vignette 
"How to use import functions".

```{r}
options(ISAnalytics.widgets = FALSE)
path_AF <- system.file("extdata", "ex_association_file.tsv",
    package = "ISAnalytics"
)

root_correct <- system.file("extdata", "fs.zip",
    package = "ISAnalytics"
)
root_correct <- unzip_file_system(root_correct, "fs")
association_file <- import_association_file(path_AF, root_correct)
```

Perform aggregation:

```{r}
aggregated_meta <- aggregate_metadata(association_file,
    grouping_keys = c(
        "SubjectID",
        "CellMarker",
        "Tissue", "TimePoint"
    ),
    import_stats = FALSE
)
```

```{r echo=FALSE}
knitr::kable(aggregated_meta)
```

As you can see there is an additional parameter you can set, `import_stats`: 
if set to `TRUE`, the function will automatically look into the file system you 
provided as root when you imported the association file and will try to locate 
first the `iss` folder for each project and then all Vispa2 "stats" files.  
Vispa2 stats contain useful information that is not included in the association 
file and is linked to the single Vispa2 run. In some cases it's useful to 
perform aggregation on those info too. If you set the parameter to `TRUE`, 
besides the columns mentioned before you will also have:

* `BARCODE_MUX` - The sum of the "BARCODE_MUX" column for the group
* `TRIMMING_FINAL_LTRLC` - The sum of the "TRIMMING_FINAL_LTRLC" column for 
the group
* `LV_MAPPED` - The sum of the "LV_MAPPED" column for the group
* `BWA_MAPPED_OVERALL` - The sum of the "BWA_MAPPED_OVERALL" column for the 
group
* `ISS_MAPPED_PP` - The sum of the "ISS_MAPPED_PP" column for the group

```{r R.options=options(ISAnalytics.widgets = FALSE)}
aggregated_meta <- aggregate_metadata(association_file,
    grouping_keys = c(
        "SubjectID", "CellMarker",
        "Tissue", "TimePoint"
    ),
    import_stats = TRUE
)
```

```{r echo=FALSE}
knitr::kable(aggregated_meta)
```

If you have the option `ISAnalytics.widgets` set to TRUE, this will produce a 
report in HTML format that tells you which stats files were imported. 
To avoid this, you can set the option to FALSE.

# Aggregation of values by key

`ISAnalytics` contains useful functions to aggregate the values contained in 
your imported matrices based on a key, aka a single column or a combination of 
columns contained in the association file that are related to the samples.

## Typical workflow

Import your association file (see previous section) and then import your 
matrices:

```{r R.options=options(ISAnalytics.widgets = FALSE)}
matrices <- import_parallel_Vispa2Matrices_auto(
    association_file = association_file, root = NULL,
    quantification_type = c("fragmentEstimate", "seqCount"),
    matrix_type = "annotated", workers = 2, patterns = NULL,
    matching_opt = "ANY"
)
```

The function `aggregate_values_by_key` can perform the aggregation both on the 
list of matrices and a single matrix.

```{r}
# Takes the whole list and produces a list in output
aggregated_matrices <- aggregate_values_by_key(matrices, association_file)

# Takes a single matrix and produces a single matrix as output
aggregated_matrices_single <- aggregate_values_by_key(matrices$seqCount, 
                                                      association_file)
```

```{r echo=FALSE}
knitr::kable(head(aggregated_matrices_single))
```


### Changing parameters to obtain different results

The function has several different parameters that have default values that 
can be changed according to user preference.

1. **Changing the `key` value**  
You can change the value of the parameter key as you see fit. This parameter 
should contain one or multiple columns of the association file that you want 
to include in the grouping when performing the aggregation. 
The default value is set to "SubjectID".

```{r}
agg1 <- aggregate_values_by_key(
    x = matrices$seqCount,
    association_file = association_file,
    key = c("SubjectID", "ProjectID")
)
```

```{r echo=FALSE}
knitr::kable(head(agg1))
```

2. **Changing the `lambda` and `args` values**  
The `lambda` parameter indicates the function to be applied to the values for 
aggregation. By default this is set to "sum", but you can change it (together 
with other parameters also), to apply other functions, both exported from 
packages or fully customized to your needs.  
The only important note is that this function should perform some kind of 
aggregation on numeric values, this means in practical terms it has to accept 
a vector of numeric/integer values as input and produce a SINGLE value as 
output. Valid options for this purpose might be: `sum`, `mean`, `median`, `min`,
`max` and so on.

```{r eval=FALSE}
agg2 <- aggregate_values_by_key(
    x = matrices$seqCount,
    association_file = association_file,
    key = "SubjectID",
    lambda = "mean"
)
```

If the function you're using accepts other parameters than the input vector, 
you can specify them in the `args` parameter as a named list:

```{r}
agg2 <- aggregate_values_by_key(
    x = matrices$seqCount,
    association_file = association_file,
    key = "SubjectID",
    lambda = "mean",
    args = list(na.rm = TRUE)
)
```

In this case the `mean` function accepts an additional parameter `na.rm` that 
we want to set to TRUE. Note that the arguments must be in a named list, where 
names are the names of the parameters to pass on to the function.

```{r echo=FALSE}
knitr::kable(head(agg2))
```

3. **Changing the `group` value**  
The `group` parameter should contain all other variables to include in the 
grouping besides `key`. By default this contains `c("chr", "integration_locus",
"strand", "GeneName", "GeneStrand")`. You can change this grouping as you see 
fit, if you don't want to add any other variable to the key, just set it to 
NULL.

```{r}
agg3 <- aggregate_values_by_key(
    x = matrices$seqCount,
    association_file = association_file,
    key = "SubjectID",
    lambda = "mean",
    args = list(na.rm = TRUE),
    group = c("chr", "integration_locus", "strand")
)
```

```{r echo=FALSE}
knitr::kable(head(agg3))
```

4. **Changing `namespace` and `env`**  
As stated before, you can specify several different functions, for them to be 
exported by the base package is not mandatory. If the function you're 
interested in applying is exported by an installed package, you should specify 
the name of the package in the `namespace` parameter.  
It is also possible to apply a fully custom function (provided it respects the 
restrictions stated before): this means you can write your own function in some
environment (most commonly the global environment) and pass it in the function 
as `lambda`. In this case you have to specify the environment in which the 
function is defined and set `namespace` as NULL.

```{r}
foo <- function(x) {
    sum(x)
}
env <- rlang::current_env()

agg4 <- aggregate_values_by_key(
    x = matrices$seqCount,
    association_file = association_file,
    key = "SubjectID",
    lambda = "foo",
    args = NULL,
    namespace = NULL,
    env = env
)
```

```{r echo=FALSE}
knitr::kable(head(agg4))
```

NOTE: when specifying `env`, note that also all other symbols (including those 
passed in `args`) will be evaluated in `env`.

```{r}
foo <- function(x, y) {
    sum(x, y)
}
c <- 2
env <- rlang::current_env()

agg4 <- aggregate_values_by_key(
    x = matrices$seqCount,
    association_file = association_file,
    key = "SubjectID",
    lambda = "foo",
    args = list(y = c),
    namespace = NULL,
    env = env
)
```

```{r echo=FALSE}
knitr::kable(head(agg4))
```

### Using the function to obtain summaries for each group based on values

As we've seen in the previous section, we can pass any lambda function to 
`aggregate_values_by_key`, provided it takes as input a vector of integer or
numeric values and produces a single value as output. This is true also for 
other objects (S3/S4 etc): you can use this function for example to produce 
summaries and include them in the tibble. A function we recommend is 
`psych::describe`.

```{r echo=FALSE}
inst <- require("psych")
```

```{r eval=inst}
agg5 <- aggregate_values_by_key(
    x = matrices$seqCount,
    association_file = association_file,
    key = "SubjectID",
    lambda = "describe",
    args = NULL,
    namespace = "psych",
    env = env
)
head(agg5)
```

To see all of the contents of the lists, use `tidyr::unnest`:

```{r eval=inst}
library(magrittr)
agg5 <- agg5 %>% tidyr::unnest(.data$Aggregated_value)
```

```{r echo=FALSE, eval=inst}
knitr::kable(head(agg5))
```

# Reproducibility

The `r Biocpkg("ISAnalytics")` package `r citep(bib[["ISAnalytics"]])`
was made possible thanks to:

* R `r citep(bib[["R"]])`
* `r Biocpkg("BiocStyle")` `r citep(bib[["BiocStyle"]])`
* `r CRANpkg("knitcitations")` `r citep(bib[["knitcitations"]])`
* `r CRANpkg("knitr")` `r citep(bib[["knitr"]])`
* `r CRANpkg("rmarkdown")` `r citep(bib[["rmarkdown"]])`
* `r CRANpkg("sessioninfo")` `r citep(bib[["sessioninfo"]])`
* `r CRANpkg("testthat")` `r citep(bib[["testthat"]])`

This package was developed using 
`r BiocStyle::Githubpkg("lcolladotor/biocthis")`.

`R` session information.

```{r reproduce3, echo=FALSE}
## Session info
library("sessioninfo")
options(width = 120)
session_info()
```


# Bibliography

This vignette was generated using `r Biocpkg("BiocStyle")` 
`r citep(bib[["BiocStyle"]])`
with `r CRANpkg("knitr")` `r citep(bib[["knitr"]])` and 
`r CRANpkg("rmarkdown")` `r citep(bib[["rmarkdown"]])` 
running behind the scenes.

Citations made with `r CRANpkg("knitcitations")` 
`r citep(bib[["knitcitations"]])`.

```{r results = "asis", echo = FALSE, warning = FALSE, message = FALSE}
## Print bibliography
bibliography()
```
