---
title: "ISAnalytics - Collision removal report"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
    theme:
      bootswatch: cosmo
      primary: "#23687A"
    navbar:
      - { title: "Website", href: "https://calabrialab.github.io/ISAnalytics/", align: right, icon: "fa-globe" }
      - { title: "GitHub", href: "https://github.com/calabrialab/ISAnalytics", align: right, icon: "fa-github" }
params:
  sample_key: null
  dynamic_cols: null
  input_info: null
  additional_info: null
  missing_info: null
  samples_info: null
  pre_info: null
  pre_stats: null
  sharing_pre: null
  coll_info: null
  post_info: null
  post_stats: null
  sharing_post: null
  summary_post: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
options(DT.warn.size = FALSE)
library(flexdashboard)
library(DT)
input_info <- params$input_info
samples_info <- params$samples_info
pre_matrix_info <- params$pre_info
pre_stats <- params$pre_stats
pre_sharing <- params$sharing_pre
coll_info <- params$coll_info
post_info <- params$post_info
post_stats <- params$post_stats
post_sharing <- params$sharing_post
post_summary <- params$summary_post
sample_key <- params$sample_key
dynamic_cols <- params$dynamic_cols
secondary_col <-  "#ff9470"
```

<!-- Page 1: info on the matrix given as input  --> 
Input info {data-orientation=rows}
===============================================================================

Row1
-------------------------------------------------------------------------------
<!-- Number of distinct integration sites found, quantification totals for
each quantification present --> 
### Distinct integration sites
```{r echo=FALSE}
valueBox(value = input_info$total_iss, color = secondary_col, icon = "fa-dna")
```

```{r echo=FALSE, eval=TRUE}
out <- purrr::map2_chr(input_info$quant_totals, names(input_info$quant_totals), function(val, name) {
  heading <- paste("###", name, "total")
  a1 <- knitr::knit_expand(text = heading)
  a2 <- knitr::knit_expand(text = paste0("`r valueBox(value = ", round(val, digits = 2), ")`"))
  paste(c(a1, a2), collapse = "\n")
})
```

`r paste(knitr::knit(text = paste(out, collapse = '\n\n')))`

Row2
-------------------------------------------------------------------------------
<!-- Just a brief explanation of what this page is about --> 
### About this page

This page contains summary information regarding the input matrix as is 
(no pre-processing steps applied)

### Independent sample key

`r sample_key`

<!-- Page 2: info on missing samples and additional samples  -->
Additional info {data-orientation=rows}
===============================================================================
```{r echo=FALSE}
venn_diag <- eulerr::venn(samples_info)
```

Row1 {data-height=500}
-------------------------------------------------------------------------------
### Overview of samples contained in matrix and association file
<!-- Venn diagram for shared samples between AF and matrix  -->
```{r}
plot(venn_diag, fills = list(fill = c("gold", "navyblue"), alpha = 0.5))
```

### About this page
<!-- Description of page content  -->
**Missing samples**  
Metadata is missing from the association file for these samples, therefore
they'll be removed prior collision processing.

**Additional samples**  
The association file contains information about these samples in the project
of interest but they're not found in the matrix provided as input.

Row2 {.tabset .tabset-fade}
-------------------------------------------------------------------------------
### Missing samples 
<!-- If present, table with missing samples  -->
```{r echo=FALSE, eval=is.null(params$missing_info), results='asis'}
cat("*Nothing to report*")
```

```{r echo=FALSE, eval=!is.null(params$missing_info)}
mi_df <- params$missing_info 
num_cols <- colnames(mi_df)[colnames(mi_df) != dynamic_cols$pcr_id]
mi_df <- mi_df |>
  dplyr::mutate(dplyr::across(
    dplyr::all_of(num_cols),
    ~ round(.x, digits = 2)
  ))
datatable(mi_df, 
              class = "stripe",
              colnames = stringr::str_replace_all(colnames(params$missing_info),
                                                  "_", " "),
              filter = "top",
          options = list(scrollY = 350))
```

### Additional samples 
<!-- If present, table with additional samples  -->
```{r echo=FALSE, eval=is.null(params$additional_info), results='asis'}
cat("*Nothing to report*")
```

```{r echo=FALSE, eval=!is.null(params$additional_info)}
ai_df <- params$additional_info 
DT::datatable(ai_df, 
              class = "stripe",
              colnames = stringr::str_replace_all(colnames(ai_df),
                                                  "_", " "),
              filter = "top",
              options = list(scrollY = 350))
```

Pre-processing {data-orientation=rows}
===============================================================================

Row1
-------------------------------------------------------------------------------
<!-- Number of distinct integration sites found, quantification totals for
each quantification present --> 

### Number of distinct integration sites

```{r}
valueBox(value = pre_matrix_info$total_iss, color = secondary_col, icon = "fa-dna")
```

```{r echo=FALSE, eval=TRUE}
out <- purrr::map2_chr(pre_matrix_info$quant_totals, names(pre_matrix_info$quant_totals), function(val, name) {
  heading <- paste("###", name, "total")
  a1 <- knitr::knit_expand(text = heading)
  a2 <- knitr::knit_expand(text = paste0("`r valueBox(value = ", round(val, digits = 2), ")`"))
  paste(c(a1, a2), collapse = "\n")
})
```

`r paste(knitr::knit(text = paste(out, collapse = '\n\n')))`

Row2
-------------------------------------------------------------------------------
<!-- Percentages on the input matrix --> 
### Number of distinct integration sites with respect to input matrix
```{r echo=FALSE}
int_ratio <- round((pre_matrix_info$total_iss / input_info$total_iss) * 100, digits = 1)
gauge(int_ratio, min = 0, max = 100, symbol = "%", gaugeSectors(
  success = c(85, 100), warning = c(40, 84), danger = c(0, 39)))
```

```{r echo=FALSE, eval=TRUE}
out <- purrr::map2_chr(pre_matrix_info$quant_totals, names(pre_matrix_info$quant_totals), function(val, name) {
  heading <- paste("###", name, "total with respect to input matrix")
  ratio <- round((val / input_info$quant_totals[[name]]) * 100, digits = 1)
  a1 <- knitr::knit_expand(text = heading)
  a2 <- knitr::knit_expand(text = paste0("`r gauge(", ratio, ", min = 0, max = 100, symbol = '%', gaugeSectors(
  success = c(85, 100), warning = c(40, 84), danger = c(0, 39)))`"))
  paste(c(a1, a2), collapse = "\n")
})
```

`r paste(knitr::knit(text = paste(out, collapse = '\n\n')))`

Row3
-------------------------------------------------------------------------------
<!-- Statistic info per pool --> 
### Per pool stats {style="width: 95% !important"}
```{r echo=FALSE}
datatable(pre_stats, class = "stripe",
              colnames = stringr::str_replace_all(colnames(pre_stats),
                                                  "_", " "),
              filter = "top", rownames = FALSE,
              options = list(scrollY = 350)) |>
formatRound(columns = colnames(pre_stats)[colnames(pre_stats) !=
                                            dynamic_cols$pool], digits = 2)
```

Row4 {data-height=800 .tabset .tabset-fade}
-------------------------------------------------------------------------------
<!-- Sharing heatmaps -->
### IS sharing: absolute values
```{r echo=FALSE, eval=is.null(pre_sharing), results='asis'}
cat("*Heatmaps saved as separate files for size issues*")
```

```{r echo=FALSE, eval=!is.null(pre_sharing)}
pre_sharing$absolute
```

### IS sharing: relative values on x
```{r echo=FALSE, eval=is.null(pre_sharing), results='asis'}
cat("*Heatmaps saved as separate files for size issues*")
```

```{r echo=FALSE, eval=!is.null(pre_sharing)}
pre_sharing$on_g1 |>
  plotly::layout(title = list(text = "IS sharing - percentages on x"))
```

### IS sharing: relative values on union
```{r echo=FALSE, eval=is.null(pre_sharing), results='asis'}
cat("*Heatmaps saved as separate files for size issues*")
```

```{r echo=FALSE, eval=!is.null(pre_sharing)}
pre_sharing$on_union |>
  plotly::layout(title = list(text = "IS sharing - percentages on union"))
```

### How to read the plots
#### How to read relative heatmaps
**Relative heatmap on x values**  
Each value in a row from left to right corresponds to the amount of shared integration sites
between x and y divided by the IS observed in the group on the x axis.

**Relative heatmap on union**  
This plot is symmetric with respect to the diagonal. Each cell contains the amount of shared
integration sites normalized by the cardinality of the union of the 2 groups x and y.

Collision info {data-orientation=columns}
===============================================================================
<!-- This page contains info about the collisions process itself -->
Col1
-------------------------------------------------------------------------------
### Collisions identified
```{r echo=FALSE}
valueBox(value = coll_info$coll_n, color = secondary_col)
```

### Collisions identified (% on pre-processing matrix)
```{r echo=FALSE}
coll_id_ratio <- round((coll_info$coll_n / pre_matrix_info$total_iss) * 100, digits = 2)
gauge(coll_id_ratio, min = 0, max = 100, symbol = "%", gaugeSectors(
  success = c(0, 30), warning = c(31, 40), danger = c(41, 100)))
```

Col2
-------------------------------------------------------------------------------
### Collisions re-assigned
```{r echo=FALSE}
valueBox(value = coll_info$reassigned, icon = "fa-recycle")
```

### Collisions re-assigned (% on pre-processing matrix)
```{r echo=FALSE}
ra_ratio_pre <- round((coll_info$reassigned / pre_matrix_info$total_iss) * 100, digits = 2)
gauge(ra_ratio_pre, min = 0, max = 100, symbol = "%")
```

### Collisions re-assigned (% on collisions)
```{r echo=FALSE}
ra_ratio_coll <- round((coll_info$reassigned / coll_info$coll_n) * 100, digits = 2)
gauge(ra_ratio_coll, min = 0, max = 100, symbol = "%")
```

Col3
-------------------------------------------------------------------------------
### Collisions removed
```{r echo=FALSE}
valueBox(value = coll_info$removed, icon = "fa-trash")
```

### Collisions removed (% on pre-processing matrix)
```{r echo=FALSE}
rem_ratio_pre <- round((coll_info$removed / pre_matrix_info$total_iss) * 100, digits = 2)
gauge(rem_ratio_pre, min = 0, max = 100, symbol = "%")
```

### Collisions removed (% on collisions)
```{r echo=FALSE}
rem_ratio_coll <- round((coll_info$removed / coll_info$coll_n) * 100, digits = 2)
gauge(rem_ratio_coll, min = 0, max = 100, symbol = "%")
```

Post-processing {data-orientation=rows}
===============================================================================
<!-- Info about the post processing matrix --> 

Row1
-------------------------------------------------------------------------------
<!-- Number of distinct integration sites found, quantification totals for
each quantification present --> 

### Number of distinct integration sites

```{r}
valueBox(value = post_info$total_iss, color = secondary_col, icon = "fa-dna")
```

```{r echo=FALSE, eval=TRUE}
out <- purrr::map2_chr(post_info$quant_totals, names(post_info$quant_totals), function(val, name) {
  heading <- paste("###", name, "total")
  a1 <- knitr::knit_expand(text = heading)
  a2 <- knitr::knit_expand(text = paste0("`r valueBox(value = ", round(val, digits = 2), ")`"))
  paste(c(a1, a2), collapse = "\n")
})
```

`r paste(knitr::knit(text = paste(out, collapse = '\n\n')))`

Row2
-------------------------------------------------------------------------------
<!-- Percentages on the pre-processing matrix --> 
### Number of distinct integration sites with respect to pre-processing matrix
```{r echo=FALSE}
int_ratio_post <- round((post_info$total_iss / pre_matrix_info$total_iss) * 100, digits = 1)
gauge(int_ratio_post, min = 0, max = 100, symbol = "%", gaugeSectors(
  success = c(85, 100), warning = c(40, 84), danger = c(0, 39)))
```

```{r echo=FALSE, eval=TRUE}
out <- purrr::map2_chr(post_info$quant_totals, names(post_info$quant_totals), function(val, name) {
  heading <- paste("###", name, "total with respect to pre-processing matrix")
  ratio <- round((val / pre_matrix_info$quant_totals[[name]]) * 100, digits = 1)
  a1 <- knitr::knit_expand(text = heading)
  a2 <- knitr::knit_expand(text = paste0("`r gauge(", ratio, ", min = 0, max = 100, symbol = '%', gaugeSectors(
  success = c(85, 100), warning = c(40, 84), danger = c(0, 39)))`"))
  paste(c(a1, a2), collapse = "\n")
})
```

`r paste(knitr::knit(text = paste(out, collapse = '\n\n')))`

Row3
-------------------------------------------------------------------------------
<!-- Statistic info per pool --> 
### Per pool stats {style="width: 95% !important"}
```{r echo=FALSE}
datatable(post_stats, class = "stripe",
              colnames = stringr::str_replace_all(colnames(post_stats),
                                                  "_", " "),
              filter = "top", rownames = FALSE,
              options = list(scrollY = 350)) |>
formatRound(columns = colnames(post_stats)[colnames(post_stats) !=
                                             dynamic_cols$pool], digits = 2)
```

Row4
-------------------------------------------------------------------------------
<!-- Per patient summary --> 
### Per patient summary {style="width: 95% !important"}
```{r echo=FALSE}
datatable(post_summary, class = "stripe",
          colnames = stringr::str_replace_all(colnames(post_summary),
                                                  "_", " "),
              filter = "top", rownames = FALSE,
              options = list(scrollY = 350))
```

Row5 {data-height=800 .tabset .tabset-fade}
-------------------------------------------------------------------------------
<!-- Sharing heatmaps -->
### IS sharing: absolute values
```{r echo=FALSE, eval=is.null(post_sharing), results='asis'}
cat("*Heatmaps saved as separate files for size issues*")
```

```{r echo=FALSE, eval=!is.null(post_sharing)}
post_sharing$absolute
```

### IS sharing: relative values on x
```{r echo=FALSE, eval=is.null(post_sharing), results='asis'}
cat("*Heatmaps saved as separate files for size issues*")
```

```{r echo=FALSE, eval=!is.null(post_sharing)}
post_sharing$on_g1 |>
  plotly::layout(title = list(text = "IS sharing - percentages on x"))
```

### IS sharing: relative values on union
```{r echo=FALSE, eval=is.null(post_sharing), results='asis'}
cat("*Heatmaps saved as separate files for size issues*")
```

```{r echo=FALSE, eval=!is.null(post_sharing)}
post_sharing$on_union |>
  plotly::layout(title = list(text = "IS sharing - percentages on union"))
```

### How to read the plots
#### How to read relative heatmaps
**Relative heatmap on x values**  
Each value in a row from left to right corresponds to the amount of shared integration sites
between x and y divided by the IS observed in the group on the x axis.

**Relative heatmap on union**  
This plot is symmetric with respect to the diagonal. Each cell contains the amount of shared
integration sites normalized by the cardinality of the union of the 2 groups x and y.

Reproducibility {data-orientation=rows}
===============================================================================

Row1 {data-height=600}
-------------------------------------------------------------------------------

### Session info {style="width: 95% !important"}
```{r attr.output='style="max-height: 580px;"'}
sessionInfo()
```

