---
output: github_document
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Installation and options

`ISAnalytics` can be installed quickly in different ways:

* You can install it via [Bioconductor](http://bioconductor.org)
* You can install it via GitHub using the package `devtools`

There are always 2 versions of the package active:

* `RELEASE` is the latest stable version
* `DEVEL` is the development version, it is the most up-to-date version where
all new features are introduced

## Installation from bioconductor

RELEASE version:
```{r eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("ISAnalytics")
```

DEVEL version:
```{r eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

# The following initializes usage of Bioc devel
BiocManager::install(version='devel')

BiocManager::install("ISAnalytics")
```

## Installation from GitHub

RELEASE:
```{r eval=FALSE}
if (!require(devtools)) {
  install.packages("devtools")
}
devtools::install_github("calabrialab/ISAnalytics",
                         ref = "RELEASE_3_13",
                         dependencies = TRUE,
                         build_vignettes = TRUE)
```

DEVEL:
```{r eval=FALSE}
if (!require(devtools)) {
  install.packages("devtools")
}
devtools::install_github("calabrialab/ISAnalytics",
                         ref = "master",
                         dependencies = TRUE,
                         build_vignettes = TRUE)
```

## Setting options

`ISAnalytics` has a verbose option that allows some functions to print 
additional information to the console while they're executing. 
To disable this feature do:

```{r OptVerbose, eval=FALSE}
# DISABLE
options("ISAnalytics.verbose" = FALSE)

# ENABLE
options("ISAnalytics.verbose" = TRUE)

```

Some functions also produce report in a user-friendly HTML format, 
to set this feature:

```{r OptWidg, eval=FALSE}
# DISABLE HTML REPORTS
options("ISAnalytics.reports" = FALSE)

# ENABLE HTML REPORTS
options("ISAnalytics.reports" = TRUE)
```

